---
title: "Experiment 2 modelling"
author: "Courtney Goodridge"
date: "21/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### TvA analysis script - statisical modelling

This script takes the magnitudedata dataset from the Experiment_2_analysis.Rmd file, and conducts a range of modelling. First I assess the diistribution at the condition and participant level, before fitting the models.

## Load packages

```{r}
if(!require(effsize)) install.packages("effsize")
library(effsize)


library(statmod)

if(!require(statmod)) install.packages("statmod")
library(statmod)

if(!require(sjstats)) install.packages("sjstats")
library(sjstats)

if(!require(car)) install.packages("car")
library(car)

if(!require(dplyr)) install.packages("dplyr")
library(dplyr)

if(!require(tidyr)) install.packages("tidyr")
library(tidyr)

if(!require(Rmisc)) install.packages("Rmisc")
library(Rmisc)

if(!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

if(!require(readr)) install.packages("readr")
library(readr)

if(!require(BayesFactor)) install.packages("BayesFactor")

if(!require(V8)) install.packages("V8")
library(V8)

if(!require(V8)) install.packages("V8")
library(V8)

if(!require(rstanarm)) install.packages("rstanarm")
library(rstanarm)

if(!require(bayestestR)) install.packages("bayestestR")
library(bayestestR)

if(!require(bayesplot)) install.packages("bayesplot")
library(bayesplot)

if(!require(emmeans)) install.packages("emmeans")
library(emmeans)

if(!require(lme4)) install.packages("lme4")
library(lme4)

if(!require(wesanderson)) install.packages("wesanderson")
library(wesanderson)

if(!require(lmerTest)) install.packages("lmerTest")
library(lmerTest)

if(!require(latex2exp)) install.packages("latex2exp")
library(latex2exp)

if(!require(ez)) install.packages("ez")
library(latex2exp)

if(!require(rstatix)) install.packages("rstatix")
library(rstatix)

if(!require(latex2exp)) install.packages("latex2exp")
library(latex2exp)

if(!require(here)) install.packages("here")
library(here)

if(!require(data.table)) install.packages("data.table")
library(data.table)

if(!require(readr)) install.packages("readr")
library(readr)
```

## Load data

```{r}
magnitudedata <- read.csv(here::here("TvA_analysis/Data/Processed/magnitudedata.csv"))

magnitudedata <- magnitudedata %>%
  dplyr::filter(EarlyResponses == FALSE)
```

## theme settings

```{r}
theme_plot <-   theme(axis.title.x = element_text(size = 15), axis.text.x = element_text(size = 12), axis.title.y = element_text(size = 15), axis.text.y = element_text(size = 12), title = element_text(size = 12), legend.title = element_text(size = 15), legend.text = element_text(size = 15), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

## Zeroing heading variable

```{r}
magnitudedata$heading <- as.numeric(magnitudedata$heading)
magnitudedata$startingpos <- as.numeric(magnitudedata$startingpos)

magnitudedata <- magnitudedata %>%
  dplyr::mutate(heading_zero = case_when(heading == 1 ~ 0,
                                         heading == 2 ~ 1))
```

This is done for easier interpretation of the main effects i.e. a heading of 1 degree is the baseline rate of error development. Thus the main effect indicates the effect of increasing this rate of error development by a factor of one. 

## Fitting model for reaction time data 

```{r}
"fitting models"
fit.gamma.rt <- glmer(FirstSteeringTime ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                      nAGQ = 0,
                      family = Gamma(link = "identity"),
              data = magnitudedata)

fit.invgauss.rt <- glmer(FirstSteeringTime ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                         nAGQ = 0,
              family = inverse.gaussian(link = "identity"),
              data = magnitudedata)

fit.gaus.rt <- lmerTest::lmer(FirstSteeringTime ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                    data = magnitudedata)

"model assessment"
AIC(fit.gamma.rt, fit.gaus.rt, fit.invgauss.rt)
summary(fit.gamma.rt)
confint(fit.gamma.rt, method = "Wald")
VarCorr(fit.gamma.rt)

"residual vs fitted"
plot(fit.gamma.rt)
```

Fitting models with Gamma, Inverse Gaussian, and normal distributions. Random effects in random slopes by heading and starting position. Intercept allowed to vary by participant. 

## Fitting model for lateral position error data 

```{r}
"fitting models"
fit.gamma.lpe <- glmer(lane_position ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                       nAGQ = 0,
              family = Gamma(link = "identity"),
              data = magnitudedata)

fit.invgauss.lpe <- glmer(lane_position ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                          nAGQ = 0,
              family = inverse.gaussian(link = "identity"),
              data = magnitudedata)

fit.gaus.lpe <- lmerTest::lmer(lane_position ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                    data = magnitudedata)

"model diagnostics"
AIC(fit.gamma.lpe, fit.gaus.lpe, fit.invgauss.lpe)
summary(fit.gaus.lpe)
confint(fit.gaus.lpe, method = "Wald")
VarCorr(fit.gaus.lpe)

"residual vs fitted"
plot(fit.gaus.lpe)
```

## Fitting model for steering magnitude data 

```{r}
"fitting models"
fit.gamma.swa <- glmer(peakSWA ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                       nAGQ = 0,
              family = Gamma(link = "identity"),
              data = magnitudedata)

fit.invgauss.swa <- glmer(peakSWA ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                          nAGQ = 0,
              family = inverse.gaussian(link = "identity"),
              data = magnitudedata)

fit.gaus.swa <- lmerTest::lmer(peakSWA ~ heading_zero * startingpos + (heading_zero + startingpos | pNum),
                    data = magnitudedata)

"model diagnostics"
AIC(fit.gamma.swa, fit.invgauss.swa)
summary(fit.gamma.swa)
confint(fit.gamma.swa, method = "Wald")
VarCorr(fit.gamma.swa)

"resdiual vs fitted plots"
plot(fit.gamma.swa)
```

## Using t-tests to investigate between-level differences for reaction time

```{r}
"2 degress of orientation offset"
between_level <- magnitudedata %>%
  dplyr::group_by(pNum, startingpos) %>%
  dplyr::summarise(rt = mean(FirstSteeringTime)) %>%
  pivot_wider(names_from = startingpos, values_from = rt) 

colnames(between_level) <- c("pNum", "pos_zero", "pos_four", "pos_eight")

between_level <- between_level %>%
  dplyr::group_by(pNum) %>%
  dplyr::mutate(zero_four = pos_zero - pos_four) %>%
  dplyr::mutate(four_eight = pos_four - pos_eight)

"0-4"
mean(between_level$zero_four)
sd(between_level$zero_four)

"4-8"
mean(between_level$four_eight)
sd(between_level$four_eight)

t.test(between_level$zero_four, between_level$four_eight, paired = TRUE)
cohen.d(between_level$zero_four, between_level$four_eight, paired = TRUE)
```

Invetsigating whether between level differences change between 0 and 4 m of starting position versus 4 m and 8 m of starting position. For reaction time, this appears to be so.

## Using t-tests to investigate between-level differences and lateral position error

```{r}
"2 degress of orientation offset"
between_level <- magnitudedata %>%
  dplyr::group_by(pNum, startingpos) %>%
  dplyr::summarise(lpe = mean(lane_position)) %>%
  pivot_wider(names_from = startingpos, values_from = lpe) 

colnames(between_level) <- c("pNum", "pos_zero", "pos_four", "pos_eight")

between_level <- between_level %>%
  dplyr::group_by(pNum) %>%
  dplyr::mutate(zero_four = pos_zero - pos_four) %>%
  dplyr::mutate(four_eight = pos_four - pos_eight)

"0-4"
mean(between_level$zero_four)
sd(between_level$zero_four)

"4-8"
mean(between_level$four_eight)
sd(between_level$four_eight)

t.test(between_level$zero_four, between_level$four_eight, paired = TRUE)
cohen.d(between_level$zero_four, between_level$four_eight, paired = TRUE)
```


Invetsigating whether between level differences change between 0 and 4 m of starting position versus 4 m and 8 m of starting position. For lateral position error, this appears to be so.

## Using t-tests to investigate between-level differences and steering wheel angle

```{r}
"2 degress of orientation offset"
between_level <- magnitudedata %>%
  dplyr::group_by(pNum, startingpos) %>%
  dplyr::summarise(swa = mean(peakSWA)) %>%
  pivot_wider(names_from = startingpos, values_from = swa) 

colnames(between_level) <- c("pNum", "pos_zero", "pos_four", "pos_eight")

between_level <- between_level %>%
  dplyr::group_by(pNum) %>%
  dplyr::mutate(zero_four = pos_zero - pos_four) %>%
  dplyr::mutate(four_eight = pos_four - pos_eight)

"0-4"
mean(between_level$zero_four)
sd(between_level$zero_four)

"4-8"
mean(between_level$four_eight)
sd(between_level$four_eight)

t.test(between_level$zero_four, between_level$four_eight, paired = TRUE)
cohen.d(between_level$zero_four, between_level$four_eight)

"1 degress of orientation offset"
between_level <- magnitudedata %>%
  dplyr::group_by(pNum, heading, startingpos) %>%
  dplyr::summarise(swa = mean(peakSWA)) %>%
  pivot_wider(names_from = startingpos, values_from = swa) 

colnames(between_level) <- c("pNum", "heading", "pos_zero", "pos_four", "pos_eight")

between_level <- between_level %>%
  dplyr::filter(heading == 1) %>%
  dplyr::group_by(pNum) %>%
  dplyr::mutate(zero_four = pos_zero - pos_four) %>%
  dplyr::mutate(four_eight = pos_four - pos_eight)

"0-4"
mean(between_level$zero_four)
sd(between_level$zero_four)

"4-8"
mean(between_level$four_eight)
sd(between_level$four_eight)

t.test(between_level$zero_four, between_level$four_eight, paired = TRUE)
cohen.d(between_level$zero_four, between_level$four_eight)
```


Invetsigating whether between level differences change between 0 and 4 m of starting position versus 4 m and 8 m of starting position. For steering wheel angle, this is not the case. Between level differences are larger between 0 and 4 versus 4 and 8 for 1 degree of orientation, however the opposite is true for 2 degrees.

This is indicaive of Accumulation, but does not fully mirror lateral position error results. This could be due to steering wheel angle being an inherently more noisy measure. 
